---
title: "Replication of Mixed Design ANOVA on Web-Based Positive Psychology Interventions"
output:
  html_document: rmarkdown::github_document
date: "2023-05-03"
---

Goals for this analysis replication:

1.  (TODO) Means and standard deviations of happiness (AHI) and depression (CES- D) measures for each intervention group on each of the six measurement occasions (pretest, posttest, 1-week follow-up, 1-month follow-up, 3-month follow-up, 6-month follow-up).
    1.  visual table to summarize the data
2.  (Done) ANOVA (mixed design analyses of variance)
    1.   (provide anova table and visuals of testing for assumptions)
3.  (Done) Visual inspection of the 95% confidence intervals lends support to the conclusions from the mixed-design analyses of variance:
    1.  visual over continuous spaced time
4.  (TODO) Multilevel linear mixed-effects models, with maximum-likelihood estimation, were used to relate longitudinal changes in AHI and CES-D scores to intervention group.

For the purpose of direct comparison and replication, we conducted similar analyses, including in the ANOVAs only those participants who completed all the study requirements (N = 72). In the next section, we report additional analyses using all enrolled participants.

## 4x6 Mixed Design ANOVA

### Load Packages
```{r, results = FALSE}
# Load the packages:

library(dplyr)
library(tidyverse)
library(ggplot2)
library(gplots)
library(psych)
library(emmeans)
library(nortest)
library(afex)
library(ggfortify)
library(ggResidpanel)

# Disable scientific notation:
options(scipen=999, digits=4)
```

### Load CSV
```{r}
#Load Files
ahi_cesd <- read_csv("data/ahi-cesd.csv")

#View Files
ahi_cesd <- as.data.frame(ahi_cesd)
head(ahi_cesd)
```

### Filter for only Non.Dropout Participants

```{r warning=FALSE}
## count unique id with all recorded ocassion c(1,2,3,4,5)

library(hash)

#create new dataframe of ahi_cesd.csv
dat <- ahi_cesd

#create a new dictionary {id:occasion}
h <- hash()

# loop through each id and add the occasion's recorded
for (i in 1:nrow(dat)) {
  key <- as.character(dat[i,'id'])
  if (!has.key(key,h)) {
    h[key] <- deframe(dat[i,'occasion'])
  } else {
    h[key] <- append(h[[key]], deframe(dat[i,'occasion']) )
  }
}

#return the count of id that completed all c(1,2,3,4,5) occasions
completed <- c()

for (i in 1:length(h)) {
  #access key
  k <- keys( h[as.character(i)] )
  
  #access values
  v <- values(h[k],USE.NAMES=FALSE, simplify=FALSE)
  v <- v[[1]]
  
  #add key to completed vector
  if (all(v == c(0,1,2,3,4,5))) {
    completed = append(completed,i)
  }
}

completed
```

### Create new df of only Non.Dropout Participants for ANOVA

```{r}
#Create new df to work with
ahi.df <- ahi_cesd %>%
  select(id, intervention, occasion, ahiTotal) %>%
  mutate(status = ifelse(id %in% completed, '1', '0'))

head(ahi.df)
```

```{r}
# Factor the predictors
#Using Signiture Strengths', "Three Good Things", "Gratitude Visit", "Recording early memories"

ahi.df$intervention <- factor(ahi.df$intervention, 
                              labels = c('Using Signiture Strengths', 
                                         "Three Good Things", "Gratitude Visit", 
                                         "Recording early memories"),
                              levels = c(1,2,3,4), 
                              ordered = FALSE)

ahi.df$occasion <- factor(ahi.df$occasion, 
                          labels = c("t0","t7","t14","t38","t98","t189"),
                          levels = c(0,1,2,3,4,5), 
                          ordered = TRUE)

ahi.df$intervention <- relevel(ahi.df$intervention, ref=4)
```

Part 2

```{r}
# Is the data balanced or unbalanced?
table(ahi.df$intervention, ahi.df$occasion)
```

Part 3

```{r warning=FALSE}
m <- aov_4(ahiTotal ~ intervention*occasion + (occasion|id), type=2, data=ahi.df)
m
summary(m)

nice(m,
     intercept=TRUE,
     correction = "none")
```

Replication of the paper's 4x6 mixed-design ANOVA. Intervention ( 4 levels) x Occasion (6 levels) on AHI scores.

Viewing the ANOVA table, intervention and the interaction involving intervention is not statistically significant. Occasion had a statistically significant effect (p = \<0.001). Potentially showing participant scores increasing over time.

```{r}
#Plot the data
emmip(m, intervention ~ occasion, CIs=TRUE,
       xlab="Time",
      ylab="ahiTotal")
```

```{r}
# Post-hoc comparisons and estimated marginal means:
posthocs <- emmeans(m, specs=pairwise ~ occasion*intervention,  model="multivariate",adjust="tukey")

posthocs$emmeans # Estimated marginal means (EMM)
posthocs$contrasts %>% 
  summary(infer = TRUE) # Post-hoc comparisons
```
