---
title: "Paper Replication of Hypothesis Testing on Web-based Positive Psychology Interventions"
output:
  html_document: rmarkdown::github_document
date: "2023-05-03"
---

Paper: Web-Based Positive Psychology Interventions: A Reexamination of Effectiveness

References:

1.  Woodworth, R. J., O'Brien-Malone, A., Diamond, M. R. and Schüz, B. (2017). Web-based positive psychology interventions: A reexamination of effectiveness. Journal of Clinical Psychology, 73(3), 218--232, DOI: <https://doi.org/10.1002/jclp.22328>
2.  Woodworth, R. J., O'Brien-Malone, A., Diamond, M. R. and Schüz, B. (2018). Data from, 'Web-based Positive Psychology Interventions: A Reexaminationof Effectiveness'. Journal of Open Psychology Data, DOI: <https://doi.org/10.5334.jopd.35>

Goal: Replicate various hypothesis testing methods for no difference between participants who completed all assessments vs participants who did not complete all assessments.

Intervention Groups:

-   1 "Using Signature Strengths"

-   2 "Three Good Things"

-   3 "Gratitude Visit"

-   4 "Recording early memories" (control condition)

# Load and Clean Data

## Creating New Variable: Status

-   Before testing, I need to load the given datafile and create a new ordinal variable of status where a participant is either a 'dropout' or 'non-dropout'. Dummy variables can also be used afterwards.

Load the data *ahi-cesd.csv* and *participant_info.csv*

```{r}
library(dplyr)
library(tidyverse)
library(sjPlot)

library(lsr)
library(psych)
library(ggplot2)

# Disable scientific notation:
options(scipen=999, digits=4)

#Load Files
ahi_cesd <- read_csv("data/ahi-cesd.csv")
participant_info <- read_csv("data/participant-info.csv")

#View Files
head(ahi_cesd)
head(participant_info)
```

Using ahi_cesd for all known AHi and CESD scores, creating a new dataframe to perform functions that will help identify each unique participant's status (dropout or non-dropout).

-   Non-dropout defined as a participant with AHI and CESD scores provided in all 6 occasions.

```{r warning=FALSE}
## count unique id with all recorded ocassion c(1,2,3,4,5)

library(hash)

#create new dataframe of ahi_cesd.csv
dat <- ahi_cesd

#create a new dictionary {id:occasion}
h <- hash()

# loop through each id and add the occasion's recorded
for (i in 1:nrow(dat)) {
  key <- as.character(dat[i,'id'])
  if (!has.key(key,h)) {
    h[key] <- deframe(dat[i,'occasion'])
  } else {
    h[key] <- append(h[[key]], deframe(dat[i,'occasion']) )
  }
}

#return the count of id that completed all c(1,2,3,4,5) occasions
completed <- c()

for (i in 1:length(h)) {
  #access key
  k <- keys( h[as.character(i)] )
  
  #access values
  v <- values(h[k],USE.NAMES=FALSE, simplify=FALSE)
  v <- v[[1]]
  
  #add key to completed vector
  if (all(v == c(0,1,2,3,4,5))) {
    completed = append(completed,i)
  }
}

```

Some pre-checks: Participants who completed at least the post-test(occasion == 1)

```{r}
# COMPLETED AT LEAST THE POSTTEST
completed_posttest <- c()

for (i in 1:length(h)) {
  #access key
  k <- keys( h[as.character(i)] )
  
  #access values
  v <- values(h[k],USE.NAMES=FALSE, simplify=FALSE)
  v <- v[[1]]
  
  #condition for any participant that completed the posttest (1)
  if (any(v %in% c(1))) {
    completed_posttest = append(completed_posttest,i)
  }
}
```

Now we have the correct counts of participants who completed all assessments, and an additional check of participants who at least completed the post-test(Occasion = 1).

```{r}
length(completed_posttest)
length(completed)
```

Great! Now that we have created vectors of participant's id under out conditions of completed or not, we can add a new column of dropout or non.dropout to the *participant_info* dataframe.

```{r}
# ADD NEW COLUMN OF DROPOUT AND NON.DROPOUT TO PARTICIPANTS 

p.dat <- data.frame(participant_info)

p.dat <- participant_info %>%
  mutate(status = ifelse(id %in% completed, '1', '0'))

#pre_test_dat contains only occurences of Occasion = 0
pre_test_dat <- ahi_cesd %>%
  filter(occasion == 0)

#add column for 
p.dat <- left_join(p.dat, pre_test_dat, by='id')

#select columns needed for the df
p.dat <- p.dat %>%
  select(id, intervention = intervention.x, sex, age, educ, income, status, ahiTotal, cesdTotal)

#ensure columns are the correct data type
p.dat$sex <- as.numeric(p.dat$sex)

head(p.dat)
```

```{r}
# ADD NEW COLUMN OF DROPOUT AND NON.DROPOUT TO AHI_CESD DF 

initial.scores.dat <- ahi_cesd %>%
  filter(occasion == 0) %>%
  mutate(status = ifelse(id %in% completed, '1', '0')) %>%
  select(id, status, occasion, elapsed.days,intervention, ahiTotal, cesdTotal)

head(initial.scores.dat)
```

Done! This table will be the main df to work on various hypothesis testing to see if there is a difference between participants that completed all assessments and participants that did not complete all assessments. This precaution is done so to move forward with the paper's intention-to-treat experiment to test ANOVA on the effects of 3 different positive psychology interventions and a control(placebo) over time.

# T-Test on Initial Happiness and Depression Scores

## Initial Happiness Scores

AHI Scores with t.test

H_0: No difference between AHI scores of dropout and non.dropouts

Checking data for violations of Assumptions of T-Test - relatively equal group sizes - relatively equal variances - interval or continuous variable is normally distributed

Paper: initial happiness scores, t(293) = 0.26, p = .79

```{r}
#Check equal variances for T-Test
var.test(ahiTotal ~ status, data = p.dat)

# Shapiro-Wilk normality test for NonDropout's ahiTotal
with(p.dat, shapiro.test(ahiTotal[status == "1"]))# p = 0.7
# Shapiro-Wilk normality test for Dropout's ahiTotal
with(p.dat, shapiro.test(ahiTotal[status == "0"])) # p = 0.9

```

Both tests of equal variance and Shapiro-Wilk normality test provided p-values that were not statistically significant to reject the null hypothesis. continuing the t-test.

```{r}
res.ahiTotal <- t.test(
  ahiTotal~status,
  data = p.dat,
  var.equal=TRUE,
  conf.level = 0.95,
  alternative = "two.sided"
)
res.ahiTotal
res.ahiTotal$p.value
```

```{r}
ggstatsplot::ggbetweenstats(data = p.dat, x = status, y = ahiTotal,
                            var.equal = TRUE,
                            effsize.type = "d", mean.ci = TRUE,
                            pairwise.comparisons = FALSE, 
                            messages = FALSE, bf.message = FALSE,
                            title = "Initial Happiness Scores", xlab = "Dropout vs Non-Dropout", ylab = "AHI Score")
```

**Initial AHI scores between dropout and non.dropout groups do not show statistically significant results (p = 0.64) , and we retain the null hypothesis that there is no difference.**

## Initial Depression Scores

CESD Scores with t.test

H_0: No difference between AHI scores of dropout and non.dropouts

Checking data for violations of Assumptions of T-Test - relatively equal group sizes - relatively equal variances - interval or continuous variable is normally distributed

Paper: initial depression scores t(293) = 0.44, p = .66

```{r}
#Check equal variances for T-Test
var.test(cesdTotal ~ status, data = p.dat) # p = 0.4

# Shapiro-Wilk normality test for NonDropout's ahiTotal
with(p.dat, shapiro.test(cesdTotal[status == "1"]))# p = 0.0003
# Shapiro-Wilk normality test for Dropout's ahiTotal
with(p.dat, shapiro.test(cesdTotal[status == "0"])) # p = 0.000000007
```

```{r}
#Since normality test failed, a wilcoxon test serves as another test of differences
wilcox.test(cesdTotal ~ status, data = p.dat, alternative = "two.sided")

```

```{r}
# Initial Depression Scores
#Paper's initial depression scores t(293) = 0.44, p = .66

res.cesdTotal <- t.test(
  cesdTotal ~ status,
  data = p.dat,
  var.equal=TRUE, 
  conf.level = 0.95,
  alternative="two.sided"
)

res.cesdTotal
```

```{r}
ggstatsplot::ggbetweenstats(data = p.dat, x = status, y = cesdTotal,
                            var.equal = TRUE,
                            effsize.type = "d", mean.ci = TRUE,
                            pairwise.comparisons = FALSE, 
                            messages = FALSE, bf.message = FALSE,
                            title = "Initial Depression Scores", xlab = "Dropout vs Non-Dropout", ylab = "CESD Score")
```

# Hypothesis Testing on Demographic Variables

## Difference in Sex with Chi-Square Test

Using Chi-Square Test to see comparison of female and male samples.

Hypothesis: There is no difference between the sample of male and female among dropout and non.dropout participants.

```{r}
#H_0: 
#  P_160 = P61 (same probability of male)
#  P_33 = P_11 (same probability of female)

# Create crosstabulations
sexfreq <- xtabs(formula = ~sex + status, data = p.dat)

chisq.test(sexfreq, correct = FALSE)    # X-squared = 0.01, df = 1, p-value = 0.92
# correct = FALSE. logical indicating whether to apply continuity correction when computing test statistic for 2x2 tables. 


```

```{r echo=FALSE}
#Cross-Tab of Sex by Status frequencies. Not enough decimal points.
sjPlot::tab_xtab(var.row = p.dat$sex, var.col = p.dat$status, title = "Differences in Sex", show.row.prc = TRUE)
```

## Intervention Groups with Chi-Square Test

X-squared(3, n = 295) = 6.15, p = 0.11

H_0: No difference between intervention groups of dropout and non.dropouts

Issues: not getting the p = 0.11....could be a round up of a round up....

```{r}
intervention.freq <- xtabs(formula = ~intervention+status, data = p.dat)
chisq.test(intervention.freq, ) # X-squared = 6.143, p = 0.1046

sjPlot::tab_xtab(
  var.row = p.dat$intervention, 
  var.col = p.dat$status, 
  statistics = "cramer",
  title = "Differences in Intervention", 
  show.exp = TRUE,
  show.row.prc = TRUE)

```

## Difference in Education Level with Wilcoxon Test

Wilcoxon W = 8154, p =.83

H_0: No difference between intervention groups of dropout and non.dropouts

```{r}
wilcox.test(x = p.dat$educ[p.dat$status == '1'], 
            y = p.dat$educ[p.dat$status == '0'],
            alternative = "two.sided")

```

## Difference in Income Level with Wilcoxon Test

Income with wilcoxon test
Wilcoxon W = 8552, p = .37

H_0: No difference between intervention groups of dropout and non.dropouts

```{r}
wilcox.test(x = p.dat$income[p.dat$status == '1'], 
            y = p.dat$income[p.dat$status == '0'],
            alternative = "two.sided")

```

## Difference in Age with T-Test

H_0: No difference between intervention groups of dropout and non.dropouts

Checking data for violations of Assumptions of T-Test
- relatively equal group sizes
- relatively equal variances
- interval or continuous variable is normally distributed

> Dropouts were on average 5.4 years younger than non-dropouts (M = 42.5 vs. 47.8), Welch t(126.1) = 3.28, p = .001.

```{r}

#Test of equal variance
var.test(age ~ status, data = p.dat)

t.test(
  age ~ status, 
  data = p.dat,
  var.equal=FALSE, 
  conf.level = 0.95
)
```
